Mimicry in Online Conversations: An Exploratory Study of Linguistic Analysis Techniques

### 2016-05-04 papers/asonam2016/paper.tex
Write up results

\documentclass[conference]{IEEEtran}
\usepackage[style=ieee, backend=biber]{biblatex}

\ifCLASSINFOpdf
	\usepackage[pdftex]{graphicx}
	\graphicspath{{img/}}
	\DeclareGraphicsExtensions{.pdf,.jpg,.png}
\else
	\usepackage[dvips]{graphicx}
 	\graphicspath{{img/}}
	\DeclareGraphicsExtensions{.pdf,.jpg,.png}
\fi

@@

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The Interactive Alignment Model is the prevailing theory of
[I probably need some help phrasing this].
In this study we investigate how several linguistic alignment techniques can be used
to detect mimicry at the lexical, syntactic and semantic levels of the
Interactive Alignment Model.

@@

\begin{enumerate}
	\item The situation model --- a representation of the situation being discussed
	\item The semantic representation --- the meaning of the utterance
	\item The syntactic representation --- the sentence structure
	\item The lexical representation --- the choice of words used
	\item The phonological representation --- ???
	\item The phonetic representation --- the sounds of an utterance
\end{enumerate}

@@

@@

\subsection{Hierarchical Alignment Model}
The Hierarchical Alignment Model (HAM) \cite{doyle2016robust} is ...
So it finds the most common words, uses these as markers, then looks at how often these markers
appear.

\subsection{Word and Document Vectors}
Word2vec \cite{mikolov2013efficient, mikolov2013distributed} is a machine learning technique that,
given a word, tries to predict the two adjacent words. Each word is then mapped to vector space
and can be directly compared. The closer the two word vectors to each other, the more semantically
related they are.

Doc2vec \cite{le2014distributed} works on the same principle as Word2vec. Here, each word is mapped to
a unique vector, and each paragraph is also mapped to a vector, which contribute to predicting the next
word in the paragraph. It can be used with not only paragraphs, but with sentences or entire documents.

[I can't find any pretrained doc2vec models trained on twitter data and don't have the data to do it
myself so this might end up relegated to future work]

@@

HAM is treated differently as it requires data to be set up in a different way.
Here we have reused the original implementation, and have formatted the data into tab separated
values files with each row being a pair of tweets, with an extra TSV file to store user information.
\footnote{[I could probably make my github repo public and put a link to the code here?]}

\section{Results}
Still not sure how to compare everything.

@@

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

>>>>

\documentclass[conference]{IEEEtran}
\usepackage[style=ieee, backend=biber]{biblatex}
\usepackage[T1]{fontenc}

\ifCLASSINFOpdf
	\usepackage[pdftex]{graphicx}
	\graphicspath{{img/}}
	\DeclareGraphicsExtensions{.pdf,.jpg,.png, .eps}
\else
	\usepackage[dvips]{graphicx}
 	\graphicspath{{img/}}
	\DeclareGraphicsExtensions{.pdf,.jpg,.png, .eps}
\fi

@@

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The Interactive Alignment Model is the prevailing theory of linguistic alignment works from a psychological standpoint.
In this study we investigate how several linguistic alignment techniques can be used to detect mimicry at the lexical, syntactic and semantic levels of the Interactive Alignment Model.

@@

\begin{enumerate}
	\item The situation model --- a representation of the situation being discussed
	\item The semantic representation --- the meaning of the utterance
	\item The syntactic representation --- the sentence structure
	\item The lexical representation --- the choice of words used
	\item The phonological representation --- the abstract representation of word sounds
	\item The phonetic representation --- the sounds of an utterance
\end{enumerate}

@@

There are many techniques that work at the lexical level, many of which can be re-purposed to be used at the syntactic level by replacing the words with phrase or dependency tree notation. One technique was used to represent each IAM layer, and one technique that works outside of the IAM.

The criteria for choosing techniques were:

\begin{itemize}
	\item Newer techniques were preferred over older ones.
	\item Availability of a source implementation or simplicity of implementation
\end{itemize}

@@

\subsection{Word Vectors}
Word2vec \cite{mikolov2013efficient, mikolov2013distributed} is a machine learning technique that, given a word, tries to predict the two adjacent words. Each word is then mapped to vector space and can be directly compared. The closer the two word vectors to each other, the more semantically related they are.

While word vectors have not been used to measure mimicry before, we posit that it may give good results for finding mimicry as it gets good results in the related field of sentiment analysis and is used to find semantic similarity of texts.

@@

\footnote{[I could probably make my github repo public and put a link to the code here?]}

\section{Results}

In general, the human classifiers agreed with each other, but the agreement is fairly weak. They most strongly agreed on syntactic mimicry, \(r > 0.28, < 0.36, p < 0.004\). Additionally there seemed to be more agreement on lexical mimicry than semantic, but p values are too high to make a definitive judgment.

The mean scores from human classifiers also show positive but weak correlation with the scores from their respective techniques. Additionally means of all human classifier scores were used as a comparison with LSM. These correlations can be seen in Table \ref{scores_techniques}.

\begin{table}[!t]
\caption{Correlations of Classifier Scores and Techniques}
\label{scores_techniques}
\centering
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{l c c}
Layer, Technique & r & p \\
\hline
Lexical, LILLA & 0.27 & 0.005 \\
Syntactic, SILLA & 0.19 & 0.057 \\
Semantic, Word2vec & 0.11 & 0.231 \\
N/A (mean scores), LSM & 0.16 & 0.111 \\
\end{tabular}
\end{table}

To observe the IAM's prediction that increased levels of mimicry on one layer results in a higher probability of mimicry in other adjacent layers, we looked at the correlations between the results of our techniques, in Table \ref{techniques}, and also between the results of the human classifiers, in Table \ref{classifications}.

\begin{table}[!t]
\caption{Correlations of Techniques}
\label{techniques}
\centering
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{l c c}
Technique & r & p \\
\hline
LILLA, SILLA & 0.26 & 0.007 \\
LILLA, Word2vec & -0.31 & 0.002 \\
SILLA, Word2vec & -0.81 & < 0.001
\end{tabular}
\end{table}

\begin{table}[!t]
\caption{Correlations of Human Classifications}
\label{classifications}
\centering
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{l c c}
Layer & r & p \\
\hline
Lexical, Syntactic & 0.57 & < 0.001 \\
Lexical, Semantic & 0.33 & < 0.001 \\
Syntactic, Semantic & 0.28 & 0.005
\end{tabular}
\end{table}

@@
